---
title: "STAT 331 Portfolio"
author: "Jonathan Flores Martinez"
format: html 
embed-resources: true
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: true
code-tools: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an B + .

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv` Example 1

```{r}
#| label: wd-1-csv-1

#Lab 3 Question 2
evaluations <- read_csv(file = here::here("Week3", "Lab3", "teacher_evals.csv"))
```

-   `csv` Example 2

```{r}
#| label: wd-1-csv-2

#Lab 2 Question 1
#Revised, previously did not have here::here components
# Revision Comments For this course, you are expected to use the here()
# function from the here package to load in your data.
# If you're not sure how this works, I'd encourage you to look over this coursework page:
surveys <- read_csv(file = here::here("Week 2", "Lab 2", "surveys.csv"))


```

-   `xlsx`

```{r}
#| label: wd-1-xlsx

# Used the PA 4 example as there was no lab example of this

library(readxl)
military <- read_xlsx("gov_spending_per_capita.xlsx",
                      sheet = "Share of Govt. spending",
                      skip  = 7,
                      n_max = 191,
                      na = c("xxx", ". .", "..")
                      )
```

**WD-2: I can select necessary columns from a dataset.**

-   Example selecting specified columns

```{r}
#| label: wd-2-ex-1

#Lab 3 Question 5
#Select cols
teacher_evals_clean <- evaluations |> 
  select(course_id,
         teacher_id,
         question_no,
         no_participants,
         resp_share,
         SET_score_avg,
         percent_failed_cur,
         academic_degree,
         seniority,
         gender) |> 
  rename(sex = gender) |> 
  filter(no_participants > 9) |> 
  mutate(teacher_id = as.character(teacher_id),
         across(
           .cols = c(question_no, no_participants, seniority),
           .fns = ~ as.integer(.x)
         ))


```

-   Example removing specified columns

```{r}
#| label: wd-2-ex-2
# Challenge 4 Open ended analysis

home_care <- c("family_based_infant",
               "family_based_toddler",
               "family_based_preschool")

center_care <- c("center_based_infant",
                 "center_based_toddler",
                 "center_based_preschool")


ca_cc_table |>
  filter( study_year == 2018) |> 
  group_by(`California Region`) |> 
    summarise(center_based_infant = median(mc_infant),
            family_based_infant = median(mfcc_infant),
            center_based_toddler = median(mc_toddler),
            family_based_toddler = median(mfcc_toddler),
            center_based_preschool = median(mc_preschool),
            family_based_preschool = median(mfcc_preschool),
            .groups = "drop"  ) |> 

  pivot_longer(cols = - `California Region`, 
               names_to = "costs") |>
  
  mutate(childcare_type = fct_collapse(.f = costs,
                          center = center_care,
                          family = home_care)) |> 

  select(- costs) |> 
  group_by(`California Region`, childcare_type) |> 
  summarise(`Mean Cost of Childcare` = mean(value), .groups = "drop") |> 
  
  pivot_wider(names_from = childcare_type, 
              values_from = `Mean Cost of Childcare`) |> 
  
  arrange(desc(center)) |> 
    rename(`Mean Cost: Center Care` = center,
           `Mean Cost: Family Care` = family)

```

-   Example selecting columns based on logical values (e.g., `starts_with()`, `ends_with()`, `contains()`, `where()`)

```{r}
#| label: wd-2-ex-3

# Lab 4 Question 7
# Revised to include starts with
# This makes the titles of the facets into mc_xx
# instead of the nice labels I had

ca_childcare |> 
  select(region, county_name, study_year,
         starts_with("mc_")) |> 
  pivot_longer(cols = c(mc_infant, mc_toddler, mc_preschool),
               names_to = "age_group",
               values_to = "median_cc_price") |>
  mutate(region = fct_reorder2(
         .f = region, .x = study_year, .y = median_cc_price),
         age_group = fct(age_group,
                         levels = c("mc_infant", "mc_toddler", "mc_preschool"))) |> 

  ggplot(mapping = aes(
    x = study_year, y = median_cc_price,
    color = region )) +
  geom_point(size = .8) +
  geom_smooth(method = loess, alpha = 0.2, linewidth = .6) +
  facet_wrap(~ age_group) +
  
  scale_x_continuous(breaks = seq(2008, 2018, 2)) +
  labs(title = "Weekly Median Price for Center-Based Childcare ($)",
       color = "California Region",
       x = "Study Year", y = "") +
  theme_bw() +
  theme(panel.spacing = unit(.5, "cm"), aspect.ratio = 1,
        axis.text.x = element_text(size = 7)
         )

```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-3-numeric-ex-1

# Lab 3 Question 5
# Filtering rows, in this case with an inequality
teacher_evals_clean <- evaluations |> 
  select(course_id, teacher_id, question_no, no_participants, resp_share,
         SET_score_avg, percent_failed_cur, academic_degree,
         seniority, gender) |> 
  rename(sex = gender) |> 
  filter(no_participants > 9) |> 
  mutate(teacher_id = as.character(teacher_id),
         across(
           .cols = c(question_no, no_participants, seniority),
           .fns = ~ as.integer(.x)
         ))

```

-   Numeric Example 2

```{r}
#| label: wd-3-numeric-ex-2

# Lab 4 Question 6
ca_childcare |> 
  select(region, study_year, mhi_2018) |>
  filter(study_year %in% c(2008, 2018)) |> 
  group_by(region, study_year) |> 
  summarise(median_household_income = median(mhi_2018), .groups = "drop") |> 
  pivot_wider(names_from = study_year,
              values_from = median_household_income) |> 
  arrange(desc(`2018`)) |> 
  rename(`California Regions` = region,
         `Median Income in 2008` = `2008`,
         `Median Income in 2018` = `2018`)

```

-   Character Example 1 (any context)

```{r}
#| label: wd-3-character

# Lab 5, Murder mystery

get_fit_now_member |> 
  left_join(get_fit_now_check_in, by = join_by(id == membership_id)) |> 
  left_join(interview, by = join_by(person_id)) |> 
  mutate(check_in_date = ymd(check_in_date)) |> 
  filter(year(check_in_date) == 2018,
         month(check_in_date) == 01,
         day(check_in_date) == 09,
         str_detect(id, pattern = "^48Z"),
         membership_status == "gold") |> 
  select(id, name, membership_status, check_in_date, transcript)

```

-   Character Example 2 (example must use functions from **stringr**)

```{r}
#| label: wd-3-string

#Lab 5, Murder mystery
# made sure that the pattern started with Annabel to account for Annabels
# with different last names

person |> 
  filter(str_detect(name, pattern = "^Annabel") &
           address_street_name == "Franklin Ave"|
           address_street_name == "Northwestern Dr" &
           address_number == max(address_number)) |> 
  left_join(interview, by = join_by(id == person_id)) |> 
  select(name, address_number, address_street_name, transcript)


```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-3-date
# Lab 5 Murder Mystery
# revised to create a date column, it makes it easier to filter for dates

get_fit_now_member |> 
  left_join(get_fit_now_check_in, by = join_by(id == membership_id)) |> 
  left_join(interview, by = join_by(person_id)) |> 
  mutate(check_in_date = ymd(check_in_date)) |> 
  filter(year(check_in_date) == 2018,
         month(check_in_date) == 01,
         day(check_in_date) == 09,
         str_detect(id, pattern = "^48Z"),
         membership_status == "gold") |> 
  select(id, name, membership_status, check_in_date, transcript)

```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-4-numeric-ex-1

#Lab 10 Question 1

mm_cm_conversion <- function(vec){
  if(!is.numeric(vec)){
    stop("Please provide a numeric input for the argument.")
  }
  vec/10
}

fish <- fish |> 
  mutate(length = mm_cm_conversion(length))

```

-   Numeric Example 2

```{r}
#| label: wd-4-numeric-ex-2

# Challenge 7 Question 4

condition_index <- function(w, len) {
  if(!is.numeric(w)){
    stop("Please provide a numeric input for the argument.")
  }
  (w * 100) / (len^3)
}

fish_index <- fish |> 
  
  filter(if_any(.cols = c(weight, length),
                .fns = ~ !is.na(.x))) |>
  mutate(condition = condition_index(weight, length))

head(fish_index)

```

-   Factor Example 1 (renaming levels)

```{r}
#| label: wd-4-factor-ex-1

# Challenge 4 
home_care <- c("family_based_infant",
               "family_based_toddler",
               "family_based_preschool")

center_care <- c("center_based_infant",
                 "center_based_toddler",
                 "center_based_preschool")


ca_cc_table |>
  filter( study_year == 2018) |> 
  group_by(`California Region`) |> 
    summarise(center_based_infant = median(mc_infant),
            family_based_infant = median(mfcc_infant),
            center_based_toddler = median(mc_toddler),
            family_based_toddler = median(mfcc_toddler),
            center_based_preschool = median(mc_preschool),
            family_based_preschool = median(mfcc_preschool),
            .groups = "drop"  ) |> 

  pivot_longer(cols = - `California Region`, 
               names_to = "costs") |>
  
  mutate(childcare_type = fct_collapse(.f = costs,
                          center = center_care,
                          family = home_care)) |> 

  select(- costs) |> 
  group_by(`California Region`, childcare_type) |> 
  summarise(`Mean Cost of Childcare` = mean(value), .groups = "drop") |> 
  
  pivot_wider(names_from = childcare_type, 
              values_from = `Mean Cost of Childcare`) |> 
  
  arrange(desc(center)) |> 
    rename(`Mean Cost: Center Care` = center,
           `Mean Cost: Family Care` = family)

```

-   Factor Example 2 (reordering levels)

```{r}
#| label: wd-4-factor-ex-2

# Lab 4 Question 7
# originally the facets were in alphbetical order.
# I wanted a specific order so I created factors to reorder the levels of the facets.
ca_childcare |> 
  select(region, county_name, study_year,
         Infant = mc_infant,
         Toddler = mc_toddler,
         Preschool = mc_preschool) |> 
  pivot_longer(cols = c(Infant, Toddler, Preschool),
               names_to = "age_group",
               values_to = "median_cc_price") |>
  mutate(region = fct_reorder2(
         .f = region, .x = study_year, .y = median_cc_price),
         age_group = fct(age_group,
                         levels = c("Infant", "Toddler", "Preschool"))) |> 

  ggplot(mapping = aes(
    x = study_year, y = median_cc_price,
    color = region )) +
  geom_point(size = .8) +
  geom_smooth(method = loess, alpha = 0.2, linewidth = .6) +
  facet_wrap(~ age_group) +
  
  scale_x_continuous(breaks = seq(2008, 2018, 2)) +
  labs(title = "Weekly Median Price for Center-Based Childcare ($)",
       color = "California Region",
       x = "Study Year", y = "") +
  theme_bw() +
  theme(panel.spacing = unit(.5, "cm"), aspect.ratio = 1,
        axis.text.x = element_text(size = 7)
         )


```

-   Character (example must use functions from **stringr**)

```{r}
#| label: wd-4-string

# Challenge 4

titlefy_mm <- function(t){
  if(!is.character(t)){
    stop("Argument is not a String")
  }
  
  t |> 
    str_replace_all(pattern = "_",
                    replacement = " ") |> 
    str_to_title() |> 
    str_replace_all(pattern = "Mm",
                   replacement = "(mm)")
}

scatter_plot <- function(df, x_axis, y_axis, category){
  t_label <- titlefy_mm(
    rlang::englue("A Scatter Plot of {{x_axis}} and {{y_axis}}"))
  x_label <- titlefy_mm(rlang::englue("{{x_axis}}"))
  y_label <- titlefy_mm(rlang::englue("{{y_axis}}"))
  c_label <- titlefy_mm(rlang::englue("{{category}}"))
  
  df |> 
    ggplot(mapping = aes(x = {{x_axis}},
                         y = {{y_axis}},
                         color = {{category}})) + 
      geom_point() +
    labs(title = t_label,
         x = x_label,
         y = y_label,
         color = c_label)
}

```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-4-date

# Lab 5
get_fit_now_member |> 
  left_join(get_fit_now_check_in, by = join_by(id == membership_id)) |> 
  left_join(interview, by = join_by(person_id)) |> 
  mutate(check_in_date = ymd(check_in_date)) |> 
  filter(year(check_in_date) == 2018,
         month(check_in_date) == 01,
         day(check_in_date) == 09,
         str_detect(id, pattern = "^48Z"),
         membership_status == "gold") |> 
  select(id, name, membership_status, check_in_date, transcript)

```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()` Example 1

```{r}
#| label: wd-5-left-ex-1

#Lab 4 Question 3
ca_childcare <- ca_childcare |> 
  left_join(y = tax_rev,
            by = join_by(county_name == entity_name, study_year == year))


```

-   `right_join()` Example 1

```{r}
#| label: wd-5-right

# Challenge 5
# revised to make a right join
# good idea as this gives the resulting data set only entries
# that exist in the interview data base which is what we care about here
person |> 
  right_join(interview, by = join_by(id == person_id)) |> 
  left_join(get_fit_now_member, by = join_by(id == person_id)) |> 
  left_join(drivers_license, by = join_by(license_id == id)) |> 
  filter(membership_status == "gold",
         str_detect(plate_number,pattern = "H42W")) |> 
  select(name.x, gender, membership_status, plate_number, transcript)

```

-   `left_join()` **or** `right_join()` Example 2

```{r}
#| label: wd-5-left-right-ex-2

#Lab 5, Murder mystery
person |> 
  filter(str_detect(name, pattern = "^Annabel") &
           address_street_name == "Franklin Ave"|
           address_street_name == "Northwestern Dr" &
           address_number == max(address_number)) |> 
  left_join(interview, by = join_by(id == person_id)) |> 
  select(name, address_number, address_street_name, transcript)


```

-   `inner_join()` Example 1

```{r}
#| label: wd-5-inner-ex-1

# Lab 5 Murder Mystery
#Revised to include an inner join.
# The inner join was used to make sure that all the rows in result
# existed in both data sets, this way the resulting data set had no nulls

#| label: following other lead
get_fit_now_member |> 
  inner_join(get_fit_now_check_in, by = join_by(id == membership_id)) |> 
  left_join(interview, by = join_by(person_id)) |> 
  filter(check_in_date == 20180109,
         str_detect(id, pattern = "^48Z"),
         membership_status == "gold") |> 
  mutate(check_in_date = ymd(check_in_date)) |> 
  select(id, name, membership_status, check_in_date, transcript)

```

-   `inner_join()` Example 2

```{r}
#| label: wd-5-inner-ex-2

#lab 5 Murder Mystery
#revised to include an inner join
# In this case the inner joins helps because the resulting data frame only
#has people who exist in the previously joined data frames
# and in the facebook event check in.

person |> 
  left_join(interview, by = join_by(id == person_id)) |> 
  left_join(drivers_license, by = join_by(license_id == id)) |> 
  inner_join(facebook_event_checkin, by = join_by(id == person_id)) |> 
  filter(car_make == "Tesla", car_model == "Model S",
         gender == "female", event_name == "SQL Symphony Concert") |> 
  mutate(date = ymd(date)) |> 
  select(name, gender, height, age, hair_color, car_make,
         car_model, event_name, date, transcript)

```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi

```

-   `anti_join()`

```{r}
#| label: wd-6-anti

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()`

```{r}
#| label: wd-7-long
# Challenge 4

home_care <- c("family_based_infant",
               "family_based_toddler",
               "family_based_preschool")

center_care <- c("center_based_infant",
                 "center_based_toddler",
                 "center_based_preschool")


ca_cc_table |>
  filter( study_year == 2018) |> 
  group_by(`California Region`) |> 
    summarise(center_based_infant = median(mc_infant),
            family_based_infant = median(mfcc_infant),
            center_based_toddler = median(mc_toddler),
            family_based_toddler = median(mfcc_toddler),
            center_based_preschool = median(mc_preschool),
            family_based_preschool = median(mfcc_preschool),
            .groups = "drop"  ) |> 

  pivot_longer(cols = - `California Region`, 
               names_to = "costs") |>
  
  mutate(childcare_type = fct_collapse(.f = costs,
                          center = center_care,
                          family = home_care)) |> 

  select(- costs) |> 
  group_by(`California Region`, childcare_type) |> 
  summarise(`Mean Cost of Childcare` = mean(value), .groups = "drop") |> 
  
  pivot_wider(names_from = childcare_type, 
              values_from = `Mean Cost of Childcare`,
              names_prefix = "mean_cost_") |> 
  
  arrange(desc(mean_cost_center))
```

-   `pivot_wider()`

```{r}
#| label: wd-7-wide

#Challenge 4 Open ended analysis
home_care <- c("family_based_infant",
               "family_based_toddler",
               "family_based_preschool")

center_care <- c("center_based_infant",
                 "center_based_toddler",
                 "center_based_preschool")

ca_cc_table |>
  filter( study_year == 2018) |> 
  group_by(`California Region`) |> 
    summarise(center_based_infant = median(mc_infant),
            family_based_infant = median(mfcc_infant),
            center_based_toddler = median(mc_toddler),
            family_based_toddler = median(mfcc_toddler),
            center_based_preschool = median(mc_preschool),
            family_based_preschool = median(mfcc_preschool),
            .groups = "drop"  ) |> 

  pivot_longer(cols = - `California Region`, 
               names_to = "costs") |>
  
  mutate(childcare_type = fct_collapse(.f = costs,
                          center = center_care,
                          family = home_care)) |> 

  select(- costs) |> 
  group_by(`California Region`, childcare_type) |> 
  summarise(`Mean Cost of Childcare` = mean(value), .groups = "drop") |> 
  
  pivot_wider(names_from = childcare_type, 
              values_from = `Mean Cost of Childcare`,
              names_prefix = "mean_cost_") |> 
  
  arrange(desc(mean_cost_center))
```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

The following assignments satisfy the above criteria:

-   Example 1: Lab 2
-   Example 2: Lab 3
-   Example 3: Lab 4
-   Example 4: Challenge 4
-   Example 5 : Lab 5

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting

```{r}
#| label: r-2-1

# Challenge 2
ggplot(data = surveys,
    mapping = aes(x = weight, y = species)) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(color = "steel blue", alpha = 0.08) +
  labs(title = "Distribution of Weight by Species Boxplot",
       subtitle = "Horizontal Orientation",
       x = "Weight (g)",
       y = "Species Name") +
  
  theme_bw() +
  theme(panel.border = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(color = "steelblue"),
        axis.ticks.y = element_line(color = "white"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

```

-   Example of **dplyr** pipeline

```{r}
#| label: r-2-2

# Lab 3 Question 5
# Revise my new line creation strategy here. I took peer feedback
# "The only feedback I have (which is a bit nitpicky)
# is that you put newlines on commas somewhat inconsistently". 
# I made it so 2 col names per row was consistent, I think that's clear
# and uncluttered
teacher_evals_clean <- evaluations |> 
  select(course_id, teacher_id,
         question_no, no_participants,
         resp_share, SET_score_avg,
         percent_failed_cur, academic_degree,
         seniority, gender) |> 
  rename(sex = gender) |> 
  filter(no_participants > 9) |> 
  mutate(teacher_id = as.character(teacher_id),
         across(
           .cols = c(question_no, no_participants, seniority),
           .fns = ~ as.integer(.x)
         ))
```

-   Example of function formatting

```{r}
#| label: r-2-3

# Lab 
# plot function

titlefy_mm <- function(t){
  if(!is.character(t)){
    stop("Argument is not a String")
  }
  t |> 
    str_replace_all(pattern = "_",
                    replacement = " ") |> 
    str_to_title() |> 
    str_replace_all(pattern = "Mm",
                   replacement = "(mm)")
}

scatter_plot <- function(df, x_axis, y_axis, category){
  t_label <- titlefy_mm(
    rlang::englue("A Scatter Plot of {{x_axis}} and {{y_axis}}"))
  x_label <- titlefy_mm(rlang::englue("{{x_axis}}"))
  y_label <- titlefy_mm(rlang::englue("{{y_axis}}"))
  c_label <- titlefy_mm(rlang::englue("{{category}}"))
  
  df |> 
    ggplot(mapping = aes(x = {{x_axis}},
                         y = {{y_axis}},
                         color = {{category}})) + 
      geom_point() +
    labs(title = t_label,
         x = x_label,
         y = y_label,
         color = c_label)
}
```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example (any context)

```{r}
#| label: r-3-example

# Lab 4 Problem 5
# Revised to include names_prefix() this way it is resistant to change
ca_childcare |> 
  select(region, study_year, mhi_2018) |>
  filter(study_year %in% c(2008, 2018)) |> 
  group_by(region, study_year) |> 
  summarise(median_household_income = median(mhi_2018), .groups = "drop") |> 
  pivot_wider(names_from = study_year,
              values_from = median_household_income,
              names_prefix = "Median Income in ") |> 
  arrange(desc(`Median Income in 2018`)) |> 
  rename(`California Regions` = region)

```

-   Example (function stops)

```{r}
#| label: r-3-function-stops

# Challenge 7
condition_index <- function(w, len) {
  if(!is.numeric(w)){
    stop("Please provide a numeric input for the argument.")
  }
  (w * 100) / (len^3)
}

fish_index <- fish |> 
  filter(if_any(.cols = c(weight, length),
                .fns = ~ !is.na(.x))) |>
  mutate(condition = condition_index(weight, length))

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   At least two numeric variables

```{r}
#| label: dvs-1-num

# Lab 4 Question 8
# revised so axis are scaled in $ values

ca_childcare |> 
  ggplot(mapping = aes(y = mc_infant, x = mhi_2018)) +
  geom_point() +
  geom_smooth(method = lm) +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_continuous(labels = scales::dollar) +
  labs(y = "Median Weekly Price for Childcare ($) - Infant",
       x = "Median Household Income 2018 ($)") 
```

-   At least one numeric variable and one categorical variable

```{r}
#| label: dvs-2-num-cat

# Lab 2 Question 6

ggplot(data = surveys,
    mapping = aes(x = species, y = weight)) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(color = "steel blue", alpha = 0.08) +
  labs(title = "Distribution of Weight by Species Boxplot",
       subtitle = "Vertical Orientation",
       x = "Species Name",
       y = "Weight (g)") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45),
        axis.ticks.length.x = unit(8, "mm"),
        axis.ticks.x = element_line(color = "white"),
        panel.border = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(color = "steelblue"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

```

-   At least two categorical variables

```{r}
#| label: dvs-2-cat

# Lab 7 Question 3
fish |> 
  mutate(trip = as.character(trip),
         trip = case_when(trip == "1" ~ "Trip 1",
                          trip == "2" ~ "Trip 2")) |> 
  filter(if_any(.cols = trip:species,
                .fns = ~ is.na(.x))) |> 
  ggplot(mapping = aes(x = year)) +
    geom_bar(mapping = aes(fill = section)) +
  facet_grid(section ~ trip, axes = "all")+
  scale_x_continuous(n.breaks = 10) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Time Distribution of Forgetting to Record Fish Weights by Trip and Site",
       subtitle = "Number of Missing Observations",
       x = "Observation Year",
       y = "") +
  theme_linedraw() +
  theme(legend.position = "none")

```

-   Dates (time series plot)

```{r}
#| label: dvs-2-date

# Lab 4 Question 7

ca_childcare |> 
  select(region, county_name, study_year,
         Infant = mc_infant,
         Toddler = mc_toddler,
         Preschool = mc_preschool) |> 
  pivot_longer(cols = c(Infant, Toddler, Preschool),
               names_to = "age_group",
               values_to = "median_cc_price") |>
  mutate(region = fct_reorder2(
         .f = region, .x = study_year, .y = median_cc_price),
         age_group = fct(age_group,
                         levels = c("Infant", "Toddler", "Preschool"))) |> 

  ggplot(mapping = aes(
    x = study_year, y = median_cc_price,
    color = region )) +
  geom_point(size = .8) +
  geom_smooth(method = loess, alpha = 0.2, linewidth = .6) +
  facet_wrap(~ age_group) +
  
  scale_x_continuous(breaks = seq(2008, 2018, 2)) +
  labs(title = "Weekly Median Price for Center-Based Childcare ($)",
       color = "California Region",
       x = "Study Year", y = "") +
  theme_bw() +
  theme(panel.spacing = unit(.5, "cm"), aspect.ratio = 1,
        axis.text.x = element_text(size = 7)
         )
  

```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can modify my plot theme to be more readable

```{r}
#| label: dvs-2-ex-1

# Challenge 2 
ggplot(data = surveys,
    mapping = aes(x = weight, y = species)) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(color = "steel blue", alpha = 0.08) +
  labs(title = "Distribution of Weight by Species Boxplot",
       subtitle = "Horizontal Orientation",
       x = "Weight (g)", y = "Species Name") +
  theme_bw() +
  theme(panel.border = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(color = "steelblue"),
        axis.ticks.y = element_line(color = "white"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

```

-   I can modify my colors to be accessible to anyone's eyes

```{r}
#| label: dvs-2-ex-2

# Lab 9 Question 8
# Revised from lab to include magenta instead of red.
# Google said it was a better alternative for color deficiencies.

fish_table |> 
  gt() |> 
  tab_header(title = "Number of Missing Values per Variable",
             subtitle = "Fish Dataset") |> 
  data_color(columns = `Number Missing`,
             palette = c("green"),
             method = "factor",
             levels = c(0), 
             na_color = "magenta"
             )

```

-   I can modify my plot titles to clearly communicate the data context

```{r}
#| label: dvs-2-ex-3

# Challenge 4
ca_cc_table |> 
  group_by(`California Region`, study_year) |> 
  summarise(`Center Based Infant` = median(mc_infant),
            `Family Based Infant` = median(mfcc_infant),
            `Center Based Toddler` = median(mc_toddler),
            `Family Based Toddler` = median(mfcc_toddler),
            `Center Based Preschool` = median(mc_preschool),
            `Family Based Preschool` = median(mfcc_preschool),
            .groups = "drop"  ) |> 
  
    pivot_longer(cols = - c(`California Region`, study_year), 
               names_to = "costs") |> 
  
  ggplot(
    mapping = aes(x = study_year, y = value, color = costs)
    ) +
    geom_jitter(alpha = 0.8 ) + 
    facet_wrap(~ `California Region`) +
  labs(x = "Year", y = "",
      color = "Types of Childcare",
      title = "Childcare Costs by Region from 2008 - 2018",
      subtitle = "Median Cost of Childcare ($)") +
  scale_y_continuous(labels = scales::label_currency(prefix = "$")) +
  
  theme_bw() +
  theme(panel.background = element_rect(fill = "gray96"))

```

-   I can modify the text in my plot to be more readable

```{r}
#| label: dvs-2-ex-4

# Lab 10
# modified the text color to implement the legend in the title
set.seed(27)
ci_dat |> 
  slice_sample(n = 40) |> 
  mutate(instance = c(1:40),
         .before = estimate) |>
  ggplot() +
    geom_segment(mapping = aes(x = conf.low,
                               xend = conf.high,
                               y =  instance,
                               yend = instance,
                               color = cover)) +
    geom_vline(xintercept = 1, linetype = "dashed") +
  labs(title = "Confidence Interval Coverage Plot",
       subtitle = "Intervals that <span style='color: #64b7f7;'>do contain</span> the real value are blue, 
      those <span style='color: #33465a;'>who don't</span> are darke blue",
       y = "",
       x = "Conficence Interval") +
  theme_linedraw() +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.subtitle =  element_markdown(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "none")

```

-   I can reorder my legend to align with the colors in my plot

```{r}
#| label: dvs-2-ex-5

# Lab 4 Question 7
ca_childcare |> 
  select(region, county_name, study_year,
         Infant = mc_infant,
         Toddler = mc_toddler,
         Preschool = mc_preschool) |> 
  pivot_longer(cols = c(Infant, Toddler, Preschool),
               names_to = "age_group",
               values_to = "median_cc_price") |>
  mutate(region = fct_reorder2(
         .f = region, .x = study_year, .y = median_cc_price),
         age_group = fct(age_group,
                         levels = c("Infant", "Toddler", "Preschool"))) |> 

  ggplot(mapping = aes(
    x = study_year, y = median_cc_price,
    color = region )) +
  geom_point(size = .8) +
  geom_smooth(method = loess, alpha = 0.2, linewidth = .6) +
  facet_wrap(~ age_group) +
  
  scale_x_continuous(breaks = seq(2008, 2018, 2)) +
  labs(title = "Weekly Median Price for Center-Based Childcare ($)",
       color = "California Region",
       x = "Study Year", y = "") +
  theme_bw() +
  theme(panel.spacing = unit(.5, "cm"), aspect.ratio = 1,
        axis.text.x = element_text(size = 7)
         )

```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors (Example 1)

```{r}
#| label: dvs-3-1-ex-1

# Challenge 2
ggplot(data = surveys,
    mapping = aes(x = weight, y = species)) +
  geom_boxplot(outliers = FALSE) +
  geom_jitter(color = "steel blue", alpha = 0.08) +
  labs(title = "Distribution of Weight by Species Boxplot",
       subtitle = "Horizontal Orientation",
       x = "Weight (g)",
       y = "Species Name") +
  
  theme_bw() +
  theme(panel.border = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(color = "steelblue"),
        axis.ticks.y = element_line(color = "white"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

```

-   I can use non-standard colors (Example 2)

```{r}
#| label: dvs-3-1-ex-2

# Lab 7 Question 3
# Feedback comment - post revision: I like your use of non-standard colors!
# The combination of color and fill is hard to see in this plot.
# I would recommend using a facet instead!

#| label: visual-of-missing-values-over-time

fish |> 
  mutate(trip = as.character(trip),
         trip = case_when(trip == "1" ~ "Trip 1",
                          trip == "2" ~ "Trip 2")) |> 
  filter(if_any(.cols = trip:species,
                .fns = ~ is.na(.x))) |> 
  ggplot(mapping = aes(x = year)) +
    geom_bar(mapping = aes(fill = section)) +
  facet_grid(section ~ trip, axes = "all")+
  scale_x_continuous(n.breaks = 10) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Time Distribution of Forgetting to Record Fish Weights by Trip and Site",
       subtitle = "Number of Missing Observations",
       x = "Observation Year",
       y = "") +
  theme_linedraw() +
  theme(legend.position = "none")

```

-   I can use annotations (e.g., `geom_text()`)

```{r}
#| label: dvs-3-2

# Lab 3 Question 7
#| layout-ncol: 2
#| layout-align: center
#| fig-cap: 
#|   - "Figure 1: Distribution of Sex"
#|   - "Figure 2: Distribution of Seniority"
#|   - "Figure 3: Distribution of Degree Held"

individuals <- teacher_evals_clean |>
  group_by(teacher_id, sex, academic_degree, seniority) |> 
  summarise(n = n(), .groups = "drop")

individuals |> 
  group_by(sex) |> 
  summarise(n = n()) |> 
  ggplot(mapping = aes(x = sex, y = n)) +
  geom_col() +
      geom_text(aes(label = n), vjust = -0.5) +
      labs(x = "Sex", y = "Count")

individuals |> 
  group_by(seniority) |> 
  summarise(n = n()) |> 
  ggplot(mapping = aes(x = seniority, y = n)) +
  geom_col() +
      geom_text(aes(label = n), vjust = -0.5) +
      labs(x = "Seniority in Years", y = "Count")

individuals |> 
  group_by(academic_degree) |> 
  summarise(n = n()) |> 
  ggplot(mapping = aes(x = academic_degree, y = n)) +
  geom_col() +
      geom_text(aes(label = n), vjust = -0.5) +
      labs(x = "Academic Degree Held", y = "Count")

individuals |> 
  filter(academic_degree == "prof") |> 
  group_by(seniority) |> 
  summarise(n = n())  

```

-   I can choose creative geometries (e.g., `geom_segment()`, `geom_ribbon)()`)

```{r}
#| label: dvs-3-3

# Lab 10 Question 13

set.seed(27)
ci_dat |> 
  slice_sample(n = 40) |> 
  mutate(instance = c(1:40),
         .before = estimate) |>
  ggplot() +
    geom_segment(mapping = aes(x = conf.low,
                               xend = conf.high,
                               y =  instance,
                               yend = instance,
                               color = cover)) +
    geom_vline(xintercept = 1, linetype = "dashed") +
  labs(title = "Confidence Interval Coverage Plot",
       subtitle = "Intervals that <span style='color: #64b7f7;'>do contain</span> the real value are blue, 
      those <span style='color: #33465a;'>who don't</span> are darke blue",
       y = "",
       x = "Conficence Interval") +
  theme_linedraw() +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.subtitle =  element_markdown(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "none")

```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()`

```{r}
#| label: dvs-4-summarize

# Challenge 4 Open Ended Analysis
ca_cc_table |> 
  group_by(`California Region`, study_year) |> 
  summarise(`Center Based Infant` = median(mc_infant),
            `Family Based Infant` = median(mfcc_infant),
            `Center Based Toddler` = median(mc_toddler),
            `Family Based Toddler` = median(mfcc_toddler),
            `Center Based Preschool` = median(mc_preschool),
            `Family Based Preschool` = median(mfcc_preschool),
            .groups = "drop"  ) |> 
  
    pivot_longer(cols = - c(`California Region`, study_year), 
               names_to = "costs") |> 
  
  ggplot(
    mapping = aes(x = study_year, y = value, color = costs)
    ) +
    geom_jitter(alpha = 0.8 ) + 
    facet_wrap(~ `California Region`) +
  labs(x = "Year", y = "",
      color = "Types of Childcare",
      title = "Childcare Costs by Region from 2008 - 2018",
      subtitle = "Median Cost of Childcare ($)") +
  scale_y_continuous(labels = scales::label_currency(prefix = "$")) +
  
  theme_bw() +
  theme(panel.background = element_rect(fill = "gray96"))

```

-   Example using `across()`

```{r}
#| label: dvs-4-across

# Lab 3 Question 6

teacher_evals_clean |> 
  summarise(across(.cols = c(course_id, teacher_id),
                   .fns = ~ n_distinct(.x)))

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r}
#| label: dvs-5-1

# Lab 3 Question 7
# Grouped by multiple variable to explore demographics
#| layout-ncol: 2
#| layout-align: center
#| fig-cap: 
#|   - "Figure 1: Distribution of Sex"
#|   - "Figure 2: Distribution of Seniority"
#|   - "Figure 3: Distribution of Degree Held"

individuals <- teacher_evals_clean |>
  group_by(teacher_id, sex, academic_degree, seniority) |> 
  summarise(n = n(), .groups = "drop")

individuals |> 
  group_by(sex) |> 
  summarise(n = n()) |> 
  ggplot(mapping = aes(x = sex, y = n)) +
  geom_col() +
      geom_text(aes(label = n), vjust = -0.5) +
      labs(x = "Sex", y = "Count")

individuals |> 
  group_by(seniority) |> 
  summarise(n = n()) |> 
  ggplot(mapping = aes(x = seniority, y = n)) +
  geom_col() +
      geom_text(aes(label = n), vjust = -0.5) +
      labs(x = "Seniority in Years", y = "Count")

individuals |> 
  group_by(academic_degree) |> 
  summarise(n = n()) |> 
  ggplot(mapping = aes(x = academic_degree, y = n)) +
  geom_col() +
      geom_text(aes(label = n), vjust = -0.5) +
      labs(x = "Academic Degree Held", y = "Count")

individuals |> 
  filter(academic_degree == "prof") |> 
  group_by(seniority) |> 
  summarise(n = n()) 

```

-   Example 2

```{r}
#| label: dvs-5-2

# Lab 3 Question 9
teacher_evals_clean |> 
  group_by(teacher_id, course_id) |> 
  summarise(number_of_instances = n()) |> 
  select(teacher_id, course_id, number_of_instances) |> 
  ungroup() |> 
  filter(number_of_instances == 9) |> 
  n_distinct()

```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   I can modify my column names to clearly communicate the data context

```{r}
#| label: dvs-6-ex-1

# Challenge 4 
# revised: added _childcare to make to add context to the column names
home_care <- c("family_based_infant",
               "family_based_toddler",
               "family_based_preschool")

center_care <- c("center_based_infant",
                 "center_based_toddler",
                 "center_based_preschool")


ca_cc_table |>
  filter( study_year == 2018) |> 
  group_by(`California Region`) |> 
    summarise(center_based_infant = median(mc_infant),
            family_based_infant = median(mfcc_infant),
            center_based_toddler = median(mc_toddler),
            family_based_toddler = median(mfcc_toddler),
            center_based_preschool = median(mc_preschool),
            family_based_preschool = median(mfcc_preschool),
            .groups = "drop"  ) |> 

  pivot_longer(cols = - `California Region`, 
               names_to = "costs") |>
  
  mutate(childcare_type = fct_collapse(.f = costs,
                          center_childcare = center_care,
                          family_childcare = home_care)) |> 

  select(- costs) |> 
  group_by(`California Region`, childcare_type) |> 
  summarise(`Mean Cost of Childcare` = mean(value), .groups = "drop") |> 
  
  pivot_wider(names_from = childcare_type, 
              values_from = `Mean Cost of Childcare`,
              names_prefix = "mean_cost_") |> 
  
  arrange(
    desc(mean_cost_center_childcare)
    )

```

-   I can modify the text in my table to be more readable (e.g., bold face for column headers)

```{r}
#| label: dvs-6-ex-2

# Lab 9 Question 6
# Bold Column Names

eval_table |> 
  gt(rowname_col = "level",
    groupname_col = "variable",
    row_group_as_column = TRUE) |> 
  fmt_percent(columns = prop) |> 
  cols_label(n = "Count of Instructors",
             prop = "Proportion by Variable Type") |> 
  tab_header(title = "Demographics for Instructors",
             subtitle = "Teacher Eval Surveys") |> 
  tab_spanner(label = "Statistics", columns = c("n", "prop")) |> 
  tab_stubhead(label = "Demographics") |> 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels())

```

-   I can arrange my table to have an intuitive ordering

```{r}
#| label: dvs-6-ex-3
# Lab 9 Question 3
# Arranged alphabetically and by variable type

survey_table |> 
  arrange(by = Variable) |> 
 gt(rowname_col = "Variable",
    groupname_col = "Data Type",
    row_group_as_column = TRUE) |> 
  tab_header(title = "Data Type for Each Variable",
             subtitle = "Rodent Information Dataset") |> 
  tab_stubhead(label = "Variable and their Data Type") |> 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_stubhead()) 

```

**DVS-7: I show creativity in my tables.**

-   I can use non-default colors

```{r}
#| label: dvs-7-ex-1

# Lab 9 Question 8
# Revised from lab to include magenta instead of red.
# Google said it was a better alternative for color deficiencies.

fish_table |> 
  gt() |> 
  tab_header(title = "Number of Missing Values per Variable",
             subtitle = "Fish Dataset") |> 
  data_color(columns = `Number Missing`,
             palette = c("green"),
             method = "factor",
             levels = c(0), 
             na_color = "magenta"
             )

```

-   I can modify the layout of my table to be more readable

```{r}
#| label: dvs-7-ex-2

# Lab 9 Question 6

eval_table |> 
  gt(rowname_col = "level",
    groupname_col = "variable",
    row_group_as_column = TRUE) |> 
  fmt_percent(columns = prop) |> 
  cols_label(n = "Count of Instructors",
             prop = "Proportion by Variable Type") |> 
  tab_header(title = "Demographics for Instructors",
             subtitle = "Teacher Eval Surveys") |> 
  tab_spanner(label = "Statistics", columns = c("n", "prop")) |> 
  tab_stubhead(label = "Demographics") |> 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels())

```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls)

```{r}
#| label: pe-1-one-call

# Lab 5 
# Revised formatting for clearer code
get_fit_now_member |> 
  left_join(get_fit_now_check_in, by = join_by(id == membership_id)) |> 
  left_join(interview, by = join_by(person_id)) |> 
  mutate(check_in_date = ymd(check_in_date)) |> 
  filter(year(check_in_date) == 2018,
         month(check_in_date) == 01,
         day(check_in_date) == 09,
         str_detect(id, pattern = "^48Z"),
         membership_status == "gold") |> 
  select(id,
         name,
         membership_status,
         check_in_date,
         transcript)


```

-   using `across()`

```{r}
#| label: pe-1-across

#Lab 3 Question 5
teacher_evals_clean <- evaluations |> 
  select(course_id,
         teacher_id,
         question_no,
         no_participants,
         resp_share,
         SET_score_avg,
         percent_failed_cur,
         academic_degree,
         seniority,
         gender) |> 
  rename(sex = gender) |> 
  filter(no_participants > 9) |> 
  mutate(teacher_id = as.character(teacher_id),
         across(
           .cols = c(question_no, no_participants, seniority),
           .fns = ~ as.integer(.x)
         ))
```

-   using functions from the `map()` family

```{r}
#| label: pe-1-map-1

# Lab 9 Question 4

cleaned_evals <- evaluations |>
  map_if(.p = is_character, as.factor) |> 
  bind_cols()
```

**PE-2: I can write functions to reduce repetition in my code.**

-   Example 1: Function that operates on vectors

```{r}
#| label: pe-2-1

# Challenge 7 Question 2

condition_index <- function(w, len) {
  if(!is.numeric(w)){
    stop("Please provide a numeric input for the argument.")
  }
  (w * 100) / (len^3)
}

fish_index <- fish |> 
  
  filter(if_any(.cols = c(weight, length),
                .fns = ~ !is.na(.x))) |>
  mutate(condition = condition_index(weight, length))

head(fish_index)

```

-   Example 2: Function that operates on data frames

```{r}
#| label: pe-2-2

# Lab 8 Question 3
titlefy_mm <- function(t){
  if(!is.character(t)){
    stop("Argument is not a String")
  }
  
  t |> 
    str_replace_all(pattern = "_",
                    replacement = " ") |> 
    str_to_title() |> 
    str_replace_all(pattern = "Mm",
                   replacement = "(mm)")
}

scatter_plot <- function(df, x_axis, y_axis, category){
  t_label <- titlefy_mm(
    rlang::englue("A Scatter Plot of {{x_axis}} and {{y_axis}}"))
  x_label <- titlefy_mm(rlang::englue("{{x_axis}}"))
  y_label <- titlefy_mm(rlang::englue("{{y_axis}}"))
  c_label <- titlefy_mm(rlang::englue("{{category}}"))
  
  df |> 
    ggplot(mapping = aes(x = {{x_axis}},
                         y = {{y_axis}},
                         color = {{category}})) + 
      geom_point() +
    labs(title = t_label,
         x = x_label,
         y = y_label,
         color = c_label)
}

```

-   Example 3: Function that operates on vectors *or* data frames

```{r}
#| label: pe-2-3

# Lab 8 Question 2

pivot_table <- function(df, col, row){
  df |>
    count({{col}}, {{row}}) |>
    pivot_wider(names_from = {{col}},
                values_from = n,
                values_fill = 0) |> 
    janitor::adorn_totals()
    
}

```

**PE-3:I can use iteration to reduce repetition in my code.**

-   using `across()`

```{r}
#| label: pe-3-across

# Challenge 7 Question 4
# if_any is equivalent to across within a filter

condition_index <- function(w, len) {
  if(!is.numeric(w)){
    stop("Please provide a numeric input for the argument.")
  }
  (w * 100) / (len^3)
}

fish_index <- fish |> 
  filter(if_any(.cols = c(weight, length),
                .fns = ~ !is.na(.x))) |>
  mutate(condition = condition_index(weight, length))

head(fish_index)

```

-   using a `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1

# Lab 9 Question 2

survey_table <- surveys |> 
  map_chr(.f = ~ class(.x)) |> 
  enframe(name = "Variable",
          value = "Data Type") 
  

kable(survey_table)

```

-   using a `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2

# Lab 9 Question 7

fish_table <- fish |> 
  map_int(.f = ~ sum(is.na(.x))) |> 
  enframe(name = "Variable", value = "Number Missing")

kable(fish_table)

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated

```{r}
#| label: pe-4-1

# Challenge 4 Open Ended Analysis
# The dollar function was superseded and replaced by the label_currency() function

ca_cc_table |> 
  group_by(`California Region`, study_year) |> 
  summarise(`Center Based Infant` = median(mc_infant),
            `Family Based Infant` = median(mfcc_infant),
            `Center Based Toddler` = median(mc_toddler),
            `Family Based Toddler` = median(mfcc_toddler),
            `Center Based Preschool` = median(mc_preschool),
            `Family Based Preschool` = median(mfcc_preschool),
            .groups = "drop"  ) |> 
  
    pivot_longer(cols = - c(`California Region`, study_year), 
               names_to = "costs") |> 
  
  ggplot(
    mapping = aes(x = study_year, y = value, color = costs)
    ) +
    geom_jitter(alpha = 0.8 ) + 
    facet_wrap(~ `California Region`) +
  labs(x = "Year", y = "",
      color = "Types of Childcare",
      title = "Childcare Costs by Region from 2008 - 2018",
      subtitle = "Median Cost of Childcare ($)") +
  scale_y_continuous(labels = scales::label_currency(prefix = "$")) +
  
  theme_bw() +
  theme(panel.background = element_rect(fill = "gray96"))


```

-   I can connect a data wrangling pipeline into a `ggplot()`

```{r}
#| label: pe-4-2

# Challenge 4 Open Ended Analysis
ca_cc_table |> 
  group_by(`California Region`, study_year) |> 
  summarise(`Center Based Infant` = median(mc_infant),
            `Family Based Infant` = median(mfcc_infant),
            `Center Based Toddler` = median(mc_toddler),
            `Family Based Toddler` = median(mfcc_toddler),
            `Center Based Preschool` = median(mc_preschool),
            `Family Based Preschool` = median(mfcc_preschool),
            .groups = "drop"  ) |> 
  
    pivot_longer(cols = - c(`California Region`, study_year), 
               names_to = "costs") |> 
  
  ggplot(
    mapping = aes(x = study_year, y = value, color = costs)
    ) +
    geom_jitter(alpha = 0.8 ) + 
    facet_wrap(~ `California Region`) +
  labs(x = "Year", y = "",
      color = "Types of Childcare",
      title = "Childcare Costs by Region from 2008 - 2018",
      subtitle = "Median Cost of Childcare ($)") +
  scale_y_continuous(labels = scales::label_currency(prefix = "$")) +
  
  theme_bw() +
  theme(panel.background = element_rect(fill = "gray96"))

```

## Data Simulation & Statisical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r}
#| label: dsm-1-1

# Lab 10 Question 8


mycifun <- function(beta0 = 2, beta1 = 1, n = 100){
  intercept = beta0
  slope = beta1
  x <- runif(n = n)
  ep <- rnorm(n = n,
              mean = 0,
              sd = 1)
  y = intercept + x * slope + ep
  
  fake_data <- enframe(x, value = "x_value") |> 
    left_join(enframe(y, value = "y_value"), join_by(name))
  
  linear_model <- lm(formula = y_value ~ x_value, data = fake_data)
  
  broom::tidy(linear_model, 
            conf.int = TRUE, 
            conf.level = 0.95) |>
    filter(term == "x_value") |> 
    select(estimate, conf.low, conf.high) |> 
    mutate(cover = case_when(1 >= conf.low & 1 <= conf.high ~ 1,
                           .default = 0))
}

set.seed(27)

ci_dat <- map(.x = 1:1000,
              .f = ~ mycifun(beta0 = 3,
                             beta1 = .5,
                             n = 100)
              )

```

-   Example 2

```{r}
#| label: dsm-1-2

# Lab 10 Question 1
randomBabies <- function(n_babies){
  baby_data <- tibble(baby = 1:n_babies,
                      house = sample(1:n_babies,
                                     size = n_babies))
  baby_data |> 
    filter(baby == house) |> 
    nrow()
  
}

set.seed(27)

results <- map_int(.x = 1:10000,
                   .f = ~ randomBabies(n_babies = 4)
                  )

```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1

```{r}
#| label: dsm-2-1

#Challenge 3 Question 3
#Chi squared test
chisq.test(x = teacher_evals_compare$SET_level,
           y = teacher_evals_compare$sen_level)

```

-   Example 2

```{r}
#| label: dsm-2-2

# Lab 4 Question 9
# Linear regression model
reg_mod1 <- lm(formula = mc_infant ~ mhi_2018,
               data = ca_childcare)
summary(reg_mod1)

```

-   Example 3

```{r}
#| label: dsm-2-3

# Lab 10 
# Questions 5 and 6

model <- lm(formula = y_value ~ x_value, data = fake_data)

tidy_summary <- broom::tidy(model, 
            conf.int = TRUE, 
            conf.level = 0.95) |> 
  filter(term == "x_value") |> 
  select(estimate, conf.low, conf.high)

tidy_summary
```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

Through my time as a student in this class I have been able to revise my thinking after getting feedback. In many of the labs the feedback helped me improve the clarity and efficiency of my code. One example that I remember is learning about more efficient ways of writing functions. Instead of using a lengthy collection of or statements one can make a c() and use the %in% operator to achieve the same goal. I took what I learned from this and many other feedback comments and used it to work on my future assignments. I learned from my mistakes and when I committed new ones I turned in revisions and learned from those as well. Overall I think revisions really helped me learn because fixing mistakes is easier that not making them and once you fix one it is easier to recognize them in the future.

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

Comments were provided on each cell with revisions. I stated what I revised and why I revised it to meet the portfolio expectations. Any feedback that inspired my revisions was also included.

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

I extended my thinking by taking what we leaned in class and expanding on that. One example that I like of this in my portfolio is the challenge 2. In class there were a few slide on themes but I looked through the documentation and used many other arguments that I think would be helpful in the context I as working it. A lot of the arguments I used, like removing grid lines along an axis were not covered in class but by exploring the documentation I was able to learn them and I did use some of those themes in future labs to increase clarity.

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

During the peer reviews I made sure to address what I though the student did well and then provide a little feedback on point of improvement that saw while admiring the good work. An example of a feedback that I would give is:

"I like how you formatted your code in a way that each pipe and + sign was the end of a line. I think you also did a good job with the commas, making sure that no line got too long. One improvement I can think of would be being more diligent when using named arguments inside of functions. An example would be using pattern =, inside of the str_detect function. Just for clarity. I think you did a great job."

Week 5 feedback

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->

During the group activities I always tried my best to be a good collaborator to my partner. I made sure to stick to my role and made an effort to do a good job at it. If as the coder, I ever got stuck, I made sure to think out loud and try to find the point of confusion. As the computer I supported my partner by asking about the end goal of what we were trying to do. I asked if there was anything missing and if there were any functions that would be helpful to our work. Communication was important not just during the class but also outside if we did not manage to finish in the provided time. On one of the activities we finished late and did not submit during class but my partner reached out via email and through that we were able to coordinate and make sure that everything was done on time.
